---
title: "Market Basket Analysis on Steam Data"
author: "Faathir Chikal Asyuraa"
date: "10/30/2019"
output:
  html_document: default
  pdf_document: default
---


```{r, error=FALSE}
#Required Packages
library(arules)
library(arulesViz)
library(tidyverse)
library(readxl)
library(knitr)
library(ggplot2)
library(plyr)
library(dplyr)
library(plotly)
library(RColorBrewer)
```

We'll analyze game transaction Data on Steam game store on 2017

```{r}
steam<-read.csv('steam.csv')
head(steam)
```

zero values on duration means that the user haven't played that game yet after purchasing it

let's do a little bit of EDA before we start the Market Basket Analysis
```{r}
str(steam)
```

It looks like user_id are recognized as int, let's change that, and take a look at the data
```{r}
steam$user_id<-as.factor(steam$user_id)
print(paste0('number of user: ',length(unique(steam$user_id))))
print(paste0('number of games: ',length(unique(steam$game))))
```

Top 10 Biggest Player Base
```{r}
steam %>% group_by(game) %>% dplyr::summarise(average_hours_spend = as.integer(mean(duration)),player=n()) %>% arrange(desc(player))
```


Top 10 Longest Played Games
```{r}
steam %>% group_by(game) %>% dplyr::summarise(average_hours_spend = as.integer(mean(duration)),player=n()) %>% arrange(desc(average_hours_spend))
```

Top 10 Longest Played Games with playerbase > 10 
```{r}
steam %>% group_by(game) %>% dplyr::summarise(average_hours_spend = as.integer(mean(duration)),player=n()) %>% filter(player>10) %>% arrange(desc(average_hours_spend)) 
```

Number of Games Owned
```{r}
ggplotly(steam %>% group_by(user_id) %>% dplyr::summarise(games_owned=n()) %>% ggplot(aes(x=games_owned))+geom_histogram(bins=100)+ggtitle('Distribution of games owned'))
```

Time Spent on Steam
```{r}
ggplotly(steam %>% group_by(user_id) %>% dplyr::summarise(time_spent=sum(duration)) %>% ggplot(aes(x=time_spent))+geom_histogram(bins=100)+ggtitle('Distribution of Time Spent on Steam (Hours)'))
```

Now, we are done with the EDA
Note that for MBA, we must shape the data so the rows are unique user, and there are a columns of list of transaction (in this case, games purchased)

```{r}
transactionData <- ddply(steam,c('user_id'), #Getting list of games per user
                       function(df1)paste(df1$game,
                       collapse = ","))
head(transactionData)
```


We no longer needed user_id column, and we need to rename the V1 column to games column
```{r}
transactionData$user_id <- NULL
colnames(transactionData) <- c('games')
```

Let's convert it to a transaction object
```{r,error=FALSE,message=FALSE,warning=FALSE}
write.csv(transactionData,"market_basket_transactions.csv", quote = FALSE, row.names = FALSE)
tr <- read.transactions('market_basket_transactions.csv', format = 'basket', sep=',')
```


Now, we'll use Apriori algorithm to determine the rules, providing minimum support and confidence, while keeping a value of maxlen


```{r}
association.rules <- apriori(tr, parameter = list(supp=0.01, conf=0.5, maxlen=10))
```

```{r}
inspect(sort(association.rules,decreasing = TRUE,
    by = "support")[1:10])
```

```{r}
plotly_arules(association.rules)
```
```{r}
plot(association.rules, method = "graph",  engine = "htmlwidget")
```

An alternative way to setup the rules, we can filter the data, so that we'll recommend a game only if the user has played it more than 2 hours 

```{r}
steam_more_than_2_hours <- steam %>% filter(duration>2)
dim(steam_more_than_2_hours)
```



We're only left with ~40000 data, 
```{r,error=FALSE,message=FALSE,warning=FALSE}
shorter_transactionData <- ddply(steam_more_than_2_hours,c('user_id'), #Getting list of games per user
                       function(df1)paste(df1$game,
                       collapse = ","))


shorter_transactionData$user_id <- NULL
colnames(shorter_transactionData) <- c('games')


write.csv(shorter_transactionData,"shorter_market_basket_transactions.csv", quote = FALSE, row.names = FALSE)
shorter_tr <- read.transactions('shorter_market_basket_transactions.csv', format = 'basket', sep=',')
summary(shorter_tr)

```


```{r}
shorter_association.rules <- apriori(shorter_tr, parameter = list(supp=0.005, conf=0.5, maxlen=10))
```
```{r}
inspect(shorter_association.rules[1:10])
```


Sorted by confidence
```{r}
inspect(sort(shorter_association.rules, decreasing = TRUE,
    by = "confidence")[1:10])
```


Sorted by support
```{r}
inspect(sort(shorter_association.rules, decreasing = TRUE,
    by = "support")[1:10])
```

```{r}
plot(shorter_association.rules, method = "graph",  engine = "htmlwidget")
```


What should you recommend if your friend plays Dota 2?

```{r}
Dota2.association.rules <- apriori(shorter_tr, parameter = list(supp=0.001, conf=0.05),appearance = list(lhs="Dota 2",default="rhs"))
plot(Dota2.association.rules, method = "graph",  engine = "htmlwidget")
```
